Formalization, Mechanization and Automation of Variants of G\"{o}del's Proof of God's Existence\thanks{This work has been supported by the German Research Foundation under grant BE2501/9-1. Bruno ????}

### 2013-08-18 Abstract/GodProof-Abstract.tex
syntax improvement

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms and definitions\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}.

\begin{align}

\emd{align}

\begin{description}
\item[(A1)] Any property necessarily implied by a positive property is positive: \\
       $\forall P \forall Q (positive P \wedge \Box \forall x (P x \Rightarrow Q x)) \Rightarrow positive Q)$
\item[(A2)] A property is positive if and only if its negation is not positive\\
       $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
\item[(A3)] the property of being God-like is positive; 
\item[(A4)] positive properties are necessarily positive; 
\item[(A5)] necessary existence is a positive property.
\item[(D1)] x is God-like if and only if x has every positive property, 
\item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
\item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
\end{description}

@@

The above variant of G\"{o}del's proof has now been 
constructed for the first-time
with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done:
\begin{itemize}
\item A detailed natural deduction proof; this proof has guided our work.
%
\item A formalization of the axioms, definitions and theorems in the TPTP THF syntax \cite{J22}.
%
\item Automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick}.
%
\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.

>>>>

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  G\"{o}del's proof \cite{Goedel1970} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
%
G\"{o}del defines God as a being who possesses all \emph{positive} properties.
He does not extensively discuss what positive properties are, 
but instead he states a few reasonable but debatable axioms that they should satisfy.
Various slightly different versions of axioms and definitions have been considered by G\"{o}del and by several philosophers who commented on his proof (e.g. \cite{Scott,Sobel,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). Our formalization employs the following axioms (A*) and definitions (D*):\footnote{
(A1), (A2), (A5), (D1), (D3) are logically equivalent to, respectively, axioms 5, 2 and 4 and definitions 1 and 3 in G\"odel's manuscript \cite{Goedel1970}. (A3) was introduced by Scott \cite{Scott} and could be derived from G\"odel's axiom 1 and (D1) in a logic with infinitary conjunction. (A4) is a weaker form of G\"odel's axiom 3. (D2) has an extra conjunct lacking in G\"odel's definition 2; this lack is believed to have been an oversight by G\"odel \cite{Hazen}.
}

\begin{align}
& \text{Any property necessarily implied by a positive property is positive:} \notag \\
& \quad \forall P \forall Q (pos\,P \wedge (\Box \forall x (P\,x \Rightarrow Q\,x)) \Rightarrow pos\,Q) \tag{A1} \\
& \text{A property is positive if and only if its negation is not positive:} \notag \\
& \quad \forall P (pos\,P \Leftrightarrow \neg (pos\,\neg P)) \tag{A2} \\
& \text{The property of being God-like is positive:} \notag \\
& \quad pos\,god \tag{A3} \\
& \text{Positive properties are necessarily positive:} \notag \\
& \quad \forall P (pos\,P \Rightarrow \Box (pos\,P))\tag{A4} \\
& \text{Necessary existence is a positive property:} \notag \\
& \quad pos\,nec\_exists \tag{A5} \\
& \text{x is God-like if and only if x has every positive property:} \notag \\
& \quad god\,x := \forall P (pos\,P \Rightarrow P x) \tag{D1} \\
& \parbox[c][2em][c]{0.8\textwidth}{A property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P:}\notag \\
& \quad ess\,P\,x := P\,x \wedge \forall Q (Q\,x \Rightarrow \Box (\forall y (P\,y \Rightarrow Q\,y))) \tag{D2} 
\end{align}

\begin{description}
\item[(A1)] Any property necessarily implied by a positive property is positive: \\
       $\forall P \forall Q ((positive P \wedge (\Box \forall x (P x \Rightarrow Q x))) \Rightarrow positive Q)$
\item[(A2)] A property is positive if and only if its negation is not positive:\\
       $\forall P ((positive P) \Leftrightarrow \neg (positive \neg P))$ 
\item[(A3)] The property of being God-like is positive:

\item[(A4)] positive properties are necessarily positive; 
\item[(A5)] necessary existence is a positive property.
\item[(D1)] x is God-like if and only if x has every positive property, 
\item[(D2)] a property P is an essence of x if and only if P is a property of x and every property Q that x has is necessarily implied by P, 
\item[(D3)] x necessarily exists if and only if every essence of x is necessarily exemplified.
\end{description}

@@

The above variant of G\"{o}del's proof has now been 
constructed for the first-time
with an unprecedent degree of detail and formality; cf.~\cite{FormalTheologyRepository}. The following has been done (and in this order):
\begin{itemize}
\item A detailed natural deduction proof; this proof has guided our work.
%
\item A formalization of the axioms, definitions and theorems in the TPTP THF syntax \cite{J22}.
%
\item Automatic verification of the consistency of the axioms and 
definitions with Nitpick \cite{Nitpick}.
%
\item Automatic demonstration of the theorems with the provers LEO-II \cite{LEO-II} and Satallax \cite{Satallax}.
